
---
title: "Replication of \"Community-Based Fact-Checking on Twitter's Birdwatch Platform\""
author: "Aisha Malik (CUNY Hunter College) & Vanessa Reino (Rider University)"
date: "6/20/2025"
output:
  html_document:
    theme: readable
    toc: true
    toc_float: true
    code_folding: hide
---


```{r setup, warning=FALSE, message=FALSE}
library(here)
library(scales)
library(tidyverse)
library(readr)
library(here)
library(parameters)

theme_set(theme_bw())

knitr::opts_chunk$set(echo = TRUE)
```

## Load community notes and rating data 
```{r load-data, warning=FALSE, message=FALSE}

birdwatch_notes <- read_tsv("C:\\Users\\ds3\\Desktop\\coursework\\week4\\project\\community-notes-2025-group-5\\vanessa\\birdwatch-public-data-2025-06-16-notes_filtered.tsv")
filtered_rating <- read_tsv("C:\\Users\\ds3\\Desktop\\coursework\\week4\\project\\community-notes-2025-group-5\\vanessa\\complete_filtered_ratings.tsv")
view(birdwatch_notes)
view(filtered_rating)

```

## Merge Data Frames
```{r join-notes-data, warning=FALSE, message=FALSE}

join_birdwatch_ratings <- left_join(birdwatch_notes, filtered_rating, by = "noteId")

view(join_birdwatch_ratings)

```

## Figure 2 - Trustworthy Sources
### Figure 2: Number of users who responded "Yes" to the question, "Did you link to sources you believe most people would consider trustworthy?".

```{r figure-2, warning=FALSE, message=FALSE}

birdwatch_notes %>%
  count(classification, trustworthySources) %>%
  ggplot(aes(x = classification, y = n, fill = factor(trustworthySources))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(labels = c(
    "MISINFORMED_OR_POTENTIALLY_MISLEADING" = "Misleading",
    "NOT_MISLEADING" = "Not Misleading"
    )) +
  labs(fill = "Trustworthy Source")

```

## Figure 3 - Number of Birdwatch Noted Per Checkbox (Misleading)
### Figure 3: Number of users who checked the answer options in response to the question, "Why do you believe this tweet may be misleading?".

```{r figure-3, warning=FALSE, message=FALSE}
birdwatch_notes %>%
summarize(across(c(9:15),sum)) %>% 
pivot_longer(everything(), names_to = "name", values_to = "total_count") %>% 
ggplot(aes(x= reorder(name, total_count), y = total_count)) + geom_bar(stat = "identity",  fill = "dark red") + coord_flip() +
scale_x_discrete(labels = c(
    "misleadingFactualError" = "Factual Error",
    "misleadingMissingImportantContext" = "Missing Important Context",
    "misleadingUnverifiedClaimAsFact" = "Unverified Claim as Fact",
    "misleadingOutdatedInformation" = "Outdated Information",
    "misleadingSatire" = "Satire",
    "misleadingOther" = "Other",
    "misleadingManipulatedMedia" = "Manipulated Media"
    )) +
xlab("Misleading Type") +
ylab("Number of Bird Watch Notes")

```


## Figure 4 - Number of Birdwatch Noted Per Checkbox ( Not Misleading)
### Figure 4: Number of users who checked the answer options in response to the question, "Why do you believe this tweet is not misleading?".

```{r figure-4, warning=FALSE, message=FALSE}
birdwatch_notes %>%
summarize(across(c(16:20),sum)) %>% 
pivot_longer(everything(), names_to = "name", values_to = "total_count") %>% 
ggplot(aes(x= reorder(name, total_count), y = total_count)) + geom_bar(stat = "identity",  fill = "dark blue") + coord_flip() +
scale_x_discrete(labels = c(
    "notMisleadingFactuallyCorrect" = "Factually Correct",
    "notMisleadingOutdatedButNotWhenWritten" = "Outdates but Not When Written",
    "notMisleadingClearlySatire" = "Clearly Satire",
    "notMisleadingOther" = "Other",
    "notMisleadingPersonalOpinion" = "Personal Opinion"
    )) +
xlab("Not Misleading Type") +
ylab("Number of Bird Watch Notes")
```


## Figure 5c - Misleading Status
### Figure 5c: The word count in text explanation of Birdwatch Notes.

```{r figure-5c, warning=FALSE, message=FALSE}

birdwatch_notes %>%
  mutate(word_count = str_count(summary, "\\w+")) %>%
  arrange(word_count) %>%
  group_by(classification, word_count) %>%   # group by both
  summarise(num_notes = n(), .groups = "drop") %>%
  arrange(classification, word_count) %>%
  group_by(classification) %>%
  mutate(
    cumulative_tot_notes = cumsum(num_notes),
    frac_notes = 1 - (cumulative_tot_notes / sum(num_notes))
  ) %>%
  ggplot(aes(x = word_count, y = 100 * frac_notes, color = classification)) +
  geom_line(size = 1) +
    labs(
    x = "Word Count",
    y = "CCDF"
  ) +
  scale_y_log10(limit = c(0.01, 100),label = comma)
```


## Figure 7a - Helpfulness Ratio
### Figure 7a: The CCDFs for helpfulness ratio. 

```{r figure-7a, warning=FALSE, message=FALSE}

join_birdwatch_ratings %>% filter(!is.na(helpful)) %>%
group_by(noteId, classification) %>% 
  summarise(total_votes = n(), helpful_votes = sum(helpful == 1), ratio_helpful = helpful_votes / total_votes, .groups = "drop")%>%
    count(classification, ratio_helpful) %>%
    arrange(ratio_helpful) %>%
    group_by(classification) %>%
  mutate(
    cumulative_tot_votes = cumsum(n),
    frac_votes = 1 - (cumulative_tot_votes / sum(n))
  ) %>%
  ggplot(aes(x = ratio_helpful , y = frac_votes, color = classification)) +
  geom_line() +
  labs(
  x = "Ratio Helpful",
  y = "CCDF(%)"
  ) +
   scale_y_log10(limit = c(0.01, 100),label = comma)



```


## Figure 7b - Total Votes
### Figure 7b: The CCDFs for total votes. 

```{r figure-7b, warning=FALSE, message=FALSE}

join_birdwatch_ratings %>% group_by(classification, noteId) %>% 
  count() %>% 
   arrange(n) %>%
  group_by(classification) %>%
  mutate(cumulative_tot_votes = cumsum(n), frac_votes = 1 - (cumulative_tot_votes / sum(n))) %>%
  ggplot(aes(x = n , y = 100 * frac_votes, color = classification)) +
  geom_line() +
  labs(
  x = "Word Count",
  y = "CCDF"
  ) +
   scale_y_log10(limit = c(0.01, 100),label = comma)


```


## Figure 8 - Number of Rating Per Checkbox (Helpful)
### Figure 8: Number of rating per checkbox users checked in reponse to the question, "What about this note was helpful to you?".
```{r figure-8, warning=FALSE, message=FALSE}

filtered_rating %>%
summarize(across(c(10:17),sum)) %>% 
pivot_longer(everything(), names_to = "name", values_to = "total_count") %>% 
ggplot(aes(x= reorder(name, total_count), y = total_count)) + geom_bar(stat = "identity",  fill = "dark blue") + coord_flip() +
scale_x_discrete(labels = c( 
  "helpfulClear" = "Clear",
  "helpfulGoodSources" = "Good Sources",
  "helpfulInformative" = "Informative",
  "helpfulEmpathetic" = "Empathetic",
  "helpfulAddressesClaim" = "Addresses Claim",
  "helpfulUniqueContext" = "Unique Context",
  "helpfulImportantContext" = "Important Context",
  "helpfulOther" = " Other"
)) +
xlab("Helpfulness Type") +
ylab("Number of Ratings")

```

## Figure 9 - Number of Ratings Per Checkbox (Unhelpful)
### Figure 9: Number of rating per checkbox users checked in reponse to the question, "Help us understand why this note was unhelpful?".
```{r figure-9, warning=FALSE, message=FALSE}
filtered_rating %>%
summarize(across(c(19:29),sum)) %>% 
pivot_longer(everything(), names_to = "name", values_to = "total_count") %>% 
ggplot(aes(x= reorder(name, total_count), y = total_count)) + geom_bar(stat = "identity",  fill = "dark red") + coord_flip() +
scale_x_discrete(labels = c( 
  "notHelpfulMissingKeyPoints" = "Missing Key Points",
  "notHelpfulSourcesMissingOrUnreliable" = " Sources Missing or Unreliable",
  "notHelpfulOpinionSpeculationOrBias" = " Opinion Speculation or Bias",
  "notHelpfulArgumentativeOrBias" = " Argumentative or Inflammatory",
  "notHelpfulIncorrect" = "Incorrect",
  "notHelpfulOther" = "Other",
  "notHelpfulOffTopic" = "Off Topic",
  "notHelpfulHardToUnderstand" = "Hard to Understand",
  "notHelpfulIrrelevantSources" = "Irrelevant Sources",
  "notHelpfulSpamHarassmentOrAbuse" = "Spam Harassment or Abuse",
  "notHelpfulOutdated" = "Outdated"
))+
xlab("Not Helpfulness Type") +
ylab("Number of Ratings")

```


# Logistic Regression
```{r clean-data, warning=FALSE, message=FALSE}

load("C:/Users/ds3/Desktop/coursework/week4/project/community-notes-2025-group-5/vanessa/data/source_tweets.Rdata", ournew_env <- new.env())
source_tweets <- ournew_env[["."]]
View(source_tweets)

birdwatch_notes_wordcount <- birdwatch_notes %>%
  mutate(word_count = str_count(summary, "\\w+")) %>%
  arrange(desc(word_count))
  view(birdwatch_notes_wordcount)

join_notes_rating_tweets <- filtered_rating %>% select(noteId, helpful)
view(join_notes_rating_tweets)

join_notes_rating_tweets <- source_tweets %>%
  inner_join(birdwatch_notes_wordcount, by = "noteId") %>%
  inner_join(join_notes_rating_tweets, by = "noteId")

view(join_notes_rating_tweets)

# Ensure classification is a factor or numeric and not introducing NAs
join_notes_rating_tweets <- join_notes_rating_tweets %>%
  mutate(
    classification = as.factor(classification),
    account_age = as.integer(interval(source_account_created_at, Sys.Date()) / years(1))
  )
view(join_notes_rating_tweets)

join_notes_rating_tweets %>% mutate(across(c(source_followers_count, source_friends_count, word_count, account_age), scale))

```

```{R model, warning=FALSE, message=FALSE}

library(broom)
library(tidyverse)

model <- glm(helpful~ classification + trustworthySources + word_count + account_age + source_friends_count + source_followers_count + source_verified,data = join_notes_rating_tweets, family = binomial)

model <- tidy(model, conf.int = TRUE, conf.level = 0.99)
summary(model)


join_notes_rating_tweets_pred <- join_notes_rating_tweets
  view(join_notes_rating_tweets_pred)

print(model)

``` 

```{r plot-model, warning=FALSE, message=FALSE}
  ggplot(model, aes(x = estimate, y = term)) +
    geom_point() +
    geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
    coord_flip() +
    scale_y_discrete(labels = c(
      "classificationNOT_MISLEADING" = "Misleading",
      "trustworthySources" = "Trustworthy Sources",
      "word_count" = "Word Count",
      "account_age" = "Account Age",
      "source_friends_count" = "Friends Count",
      "source_followers_count" = "Followers Count",
      "source_verifiedTRUE" = "Verified"
    )) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

```{r figure-10, warning=FALSE, message=FALSE}

library(MASS)

model_data <- join_notes_rating_tweets_pred %>% group_by(noteId) %>% mutate(helpful_votes = sum(helpful, na.rm = TRUE),
  total_votes = n(), helpfulness_ratio = helpful_votes / total_votes) %>%
  ungroup() %>% filter(!is.na(helpfulness_ratio),is.finite(helpfulness_ratio))
view(model_data)

lm_helpfullness_ratio <- lm( helpfulness_ratio~ classification + trustworthySources + word_count + 
                       account_age + source_friends_count + source_followers_count + source_verified,
                     data = model_data )

helpfullness_ratio_summary <- tidy(lm_helpfullness_ratio, conf.int = TRUE, conf.level = 0.99)
print(helpfullness_ratio_summary)

model_data <- model_data %>%
  mutate(source_verified = as.numeric(source_verified), 
  classification = as.numeric(classification),
  trustworthySources = as.numeric(trustworthySources)) %>%
  mutate(across(c(classification, trustworthySources, word_count, source_friends_count, source_followers_count, source_verified, account_age), scale, .names = "{.col}_z"))

model_votes <- glm.nb(total_votes ~ classification_z + trustworthySources_z + word_count_z + source_friends_count_z +
  source_followers_count_z + source_verified_z + account_age_z, data = model_data)

model_summary_votes <- tidy(model_votes, conf.int = TRUE, conf.level = 0.99)
print(model_summary_votes)

helpfullness_ratio_summary <- helpfullness_ratio_summary %>% mutate(DV = "Helpfulness Ratio")
model_summary_votes <- model_summary_votes |> mutate(DV = "Total Votes")

combined_summary <- bind_rows(helpfullness_ratio_summary, model_summary_votes)

combined_summary <- combined_summary |>
  filter(term != "(Intercept)") |>
  mutate(
    term = case_when(
      grepl("classification", term) ~ "Misleading",
      grepl("trustworthySources", term) ~ "Trustworthy Sources",
      grepl("word_count", term) ~ "Word Count",
      grepl("source_friends_count", term) ~ "Source Friends Count",
      grepl("source_followers_count", term) ~ "Source Followers Count",
      grepl("source_verified", term) ~ "Source Verified",
      grepl("account_age", term) ~ "Source Account Age",
      TRUE ~ term
    )
  )

  ggplot(combined_summary, aes(x = estimate, y = term, color = DV)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, position = position_dodge(width = 0.5)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  labs(title = "Figure 10: Regression Results", x = "Standardized Coefficient Estimate", y = NULL, color = "Dependent Variable") +
  theme_minimal() +
  coord_flip() +
  scale_y_discrete(labels = c(
  "classificationNOT_MISLEADING" = "Misleading",
  "trustworthySources" = "Trustworthy Sources",
  "word_count" = "Word Count",
  "account_age" = "Account Age",
  "source_friends_count" = "Friends Count",
  "source_followers_count" = "Followers Count",
  "source_verifiedTRUE" = "Verified"
    )) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

